---
title: "Final Project Visualizations 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(rio)
library(here)
library(gghighlight)
library(colorblindr)
library(janitor)
library(here)
library(zipcode)
```

```{r data import}
d <- import(here("data", "school_survey.sav"), setclass = "tbl_df")
```


```{r data manip vis 1}
d1<- d %>%
  select("school_county", 93:140) %>%
  mutate(school_county = tolower(school_county)) %>%
  rename(county = school_county) %>%
  factorize()

d2 <- d1 %>%
  group_by(county) %>%
  summarize(mean = mean(prepare))


map <- ggplot2::map_data("county") %>%
  filter(region == "oregon") %>%
  rename(county = subregion)


d1map <- left_join(map, d2)

d3 <- d1 %>%
  mutate(information_assistance_n = as.numeric(information_assistance_n)) %>%
  filter(!is.na(information_assistance_n)) %>%
  group_by(county) %>%
  summarize(mean = mean(information_assistance_n))
d3
d2map <- left_join(map, d3)
```



```{r vis 1}
ggplot() + geom_polygon(data = d1map, aes(x=long, y = lat, fill = mean, group = group)) + 
  coord_fixed(1.3) + 
  scale_fill_viridis_c()+
  theme_void() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(title = "Oregon Public Schools' Suicide Preparedness Levels", subtitle = "Average Preparedness by County", x = NULL, y = NULL, fill = "Preparedness
       Score", caption = "*Score represents a composite preparedness rating ranging from 1-5")
#would really like a second map that breaks it down by zipcode, I attempted with the package zipcode, but it just represents zipcodes with points not actual mapping. 

```

Great plot! I had used a lot of bulky code to get ride of my axes and totally forgot that this can be done in a simpler way using theme(axis.text.x = element_blank(),axis.text.y = element_blank()). and x = NULL, y = NULL. Overall, I'm impressed by your use in mapping data to the counties in Oregon!

Suggestions:
(1) For people not from Oregon, it might be nice to note the extreme counties (low and high) by name using geom_text,etc.
(2) I was curious exactly what you are measuring with suicide preparedness levels. A note about what that means in the caption describing the composite score might be helpful.


```{r vis 2}
d1 <- d1 %>%
  mutate(school_type_n = recode(school_type_n, "Middle School" = "Middle", "High School" = "High")) %>%
  mutate(region_class_n = factor(region_class_n, levels = c("Urban", "Rural", "Frontier")))
d1

d1 %>%
  filter(!is.na(school_type_n)) %>%
  ggplot(aes(school_type_n, prepare, fill = region_class_n)) + #I like the additional aesthetic of fill by region class, but can be dissuaded if it is distracting. 
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  geom_bar(stat = "summary", fun.y = "mean", alpha = 0.7) +
  #geom_errorbar(aes(school_type_n, ymin = prepare - sd, ymax = prepare + sd), width=0.2, size=1, color="blue") +
  facet_wrap(~region_class_n) +
  labs(title = "School Preparedness by Region and Level", x = "School Level", y = "Preparedness Score")
  
```
I think this plot does a great job at showing how preparedness increases as people get farther along in the school system for urban and rural communities, but that there is actually a decrease in frontier communities. You're plot is also very clear in showing there is a decrease in preparedness from urban to rural communities and then from rural to frontier. I also really like the colors you chose!

Suggestions:
(1) Is this still referring to the same "preparedness" variable as the last plot and the same data? Would be helpful to maybe make your title more specific and/or maybe add a descriptive subtitle (and also mention that this is Oregon again). This applies to the following plots as well.
(2) Along the same lines, what is "Frontier" referring to? I'm assuming rural areas near the mountains, but I'm suprised that's a different group.


```{r vis 3}
ggplot() + geom_polygon(data = d2map, aes(x=long, y = lat, fill = mean, group = group)) + 
  coord_fixed(1.3) + 
  scale_fill_viridis_c() +
 theme_void() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(title = "Oregon Public Schools' Level of Interest in Recieving Support", subtitle = "Average Interest by County", x = NULL, y = NULL, fill = "Interest", caption = "*Interest was measured on a 1-3 likert-type scale")
```

Interesting that the cities along the I-5 aren't the highest in being interested to receive support! As with the last plot about the state, I think this does a good job of displaying the data and I'm impressed by your plot. Good job!

Suggestions:
(1) I wonder if it would be helpful to clearly define the county borders with white (same in your other plot), since a lot of the colors side by side are very similar and sometimes it is difficult to see where one stops and the other starts. Labels of high/low counties might also help.
(2) Might be helpful to pull people's attention to the highest and lowest counties also, since there is a lot of variation going on in this map.



```{r vis 4}
d4 <- d1 %>%
  mutate(information_assistance_n = as.numeric(information_assistance_n)) 

d4 %>%
  filter(!is.na(school_type_n)) %>%
  ggplot(aes(school_type_n, information_assistance_n, fill = region_class_n)) + #I like the additional aesthetic of fill by region class, but can be dissuaded if it is distracting. 
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  geom_bar(stat = "summary", fun.y = "mean", alpha = 0.7) +
  #geom_errorbar(aes(school_type_n, ymin = prepare - sd, ymax = prepare + sd), width=0.2, size=1, color="blue") +
  facet_wrap(~region_class_n) +
  labs(title = "Interest in Receiving Support", x = "School Level", y = "Interest", caption = "Interest measured on a 1-3 likert-type rating scale")
```

I like that you maintained the same colors for the different regions in this plot as you did in the second plot! Again, I do think it shows the distinctive differences between regions and age groups in receiving support for their students. I'm surprised urban high schools have less interest than rural and frontier schools.

Suggestions:
(1) My same suggestions about specific title of what is being assessed and if this is about Oregon again applies to this plot.

Overall, very good job and interesting project!



